## Materials and Methods

### RNA-Seq Data Processing

Raw sequencing reads in FASTQ format were processed using the RNAdetector pipeline (v`r params$pipeline_version`) [@rnadetector_cite]. All analyses were conducted using R [@r_cite] with packages from the Bioconductor project [@bioconductor_cite].

`r if(params$trim_enabled) paste0("#### Quality Trimming\n\nQuality trimming and adapter removal were performed using Trim Galore v0.6.6 [@trim_galore_cite], which wraps Cutadapt [@cutadapt_cite] and FastQC [@fastqc_cite]. Reads were trimmed with a minimum Phred quality score threshold of ", params$trim_quality, " and a minimum read length of ", params$trim_length, " bp. Adapter sequences were auto-detected and removed.")`

`r if(!params$trim_enabled) "No quality trimming was applied to the raw reads prior to alignment."`

### Read Alignment

```{r alignment_methods, echo=TRUE, results='asis'}
# Generate alignment methods text based on the algorithm used
if (params$algorithm == "star") {
  cat("Reads were aligned to the ", params$genome_name, " reference genome using STAR v2.7.9a [@star_cite], a splice-aware aligner optimized for RNA-Seq data. ",
      "STAR was run in two-pass mode with ", params$threads, " threads. ",
      "The genome index was generated using the reference annotation ", params$annotation_name, ". ",
      "Default STAR parameters were used unless otherwise noted, including a maximum of 10 multi-mapping locations per read. ",
      "Aligned reads were output in BAM format, sorted by coordinate.\n", sep = "")
} else if (params$algorithm == "hisat2") {
  cat("Reads were aligned to the ", params$genome_name, " reference genome using HISAT2 v2.2.1 [@hisat2_cite]. ",
      "HISAT2 employs a graph-based alignment approach using a Hierarchical Graph FM index. ",
      "The alignment was performed with ", params$threads, " threads using the reference annotation ", params$annotation_name, ". ",
      "Known splice sites were extracted from the annotation file and provided to HISAT2 to guide spliced alignment. ",
      "Output alignments were converted to sorted BAM format using SAMtools [@samtools_cite].\n", sep = "")
} else if (params$algorithm == "salmon") {
  cat("Transcript-level quantification was performed using Salmon v1.5.2 [@salmon_cite] in quasi-mapping mode (pseudo-alignment). ",
      "Salmon was run with ", params$threads, " threads against a transcriptome index built from the ", params$genome_name, " reference genome ",
      "and the ", params$annotation_name, " annotation. ",
      "Salmon employs a lightweight mapping algorithm that does not produce traditional genome alignments, ",
      "instead directly estimating transcript abundances using an expectation-maximization algorithm with GC bias and sequence-specific bias correction enabled.\n", sep = "")
} else if (params$algorithm == "bwa") {
  cat("Reads were aligned to the ", params$genome_name, " reference genome using BWA-MEM v0.7.17 [@bwa_cite]. ",
      "BWA-MEM was selected for its suitability in aligning reads with potential back-splice junctions for circular RNA detection. ",
      "The alignment was performed with ", params$threads, " threads. ",
      "Output alignments were converted to sorted BAM format using SAMtools [@samtools_cite].\n", sep = "")
}
```

### Read Quantification

```{r counting_methods, echo=TRUE, results='asis'}
# Generate quantification methods text based on the counting algorithm used
if (params$counting_algorithm == "featurecounts") {
  cat("Gene-level read counts were obtained using featureCounts from the Subread package v2.0.3 [@featurecounts_cite; @subread_cite]. ",
      "Reads were assigned to genomic features based on the ", params$annotation_name, " annotation in GTF format. ",
      "featureCounts was run with ", params$threads, " threads. ",
      "Multi-mapping reads were not counted, and reads were required to have a minimum mapping quality of 10. ",
      "Only reads overlapping exonic regions were counted, with a minimum overlap of 1 bp.\n", sep = "")
} else if (params$counting_algorithm == "htseq") {
  cat("Gene-level read counts were obtained using HTSeq v0.13.5 [@htseq_cite]. ",
      "Reads were assigned to features based on the ", params$annotation_name, " annotation. ",
      "The union mode was used for handling reads overlapping multiple features. ",
      "Reads with a mapping quality below 10 were excluded from counting.\n", sep = "")
} else if (params$counting_algorithm == "salmon") {
  cat("Transcript-level abundance estimates were obtained directly from Salmon [@salmon_cite] as described above. ",
      "Transcript-per-million (TPM) and estimated read counts were computed using Salmon's variational Bayesian EM algorithm. ",
      "Gene-level counts were derived by aggregating transcript-level estimates.\n", sep = "")
} else if (params$counting_algorithm == "stringtie") {
  cat("Transcript assembly and quantification were performed using StringTie v2.1.7 [@stringtie_cite]. ",
      "StringTie was run with the ", params$annotation_name, " reference annotation to guide transcript assembly. ",
      "The algorithm estimated transcript abundances as FPKM and TPM values, ",
      "and gene-level count matrices were extracted for downstream differential expression analysis using the prepDE.py script provided with StringTie.\n", sep = "")
}
```

`r if(exists("params") && !is.null(params$de_algorithm) && params$de_algorithm != "none") "### Differential Expression Analysis"`

```{r de_methods, echo=TRUE, results='asis'}
# Generate differential expression methods text based on DE algorithm
if (!is.null(params$de_algorithm) && params$de_algorithm != "none") {
  if (params$de_algorithm == "deseq2") {
    cat("Differential expression analysis was performed using DESeq2 v1.34.0 [@deseq2_cite]. ",
        "Raw read counts were normalized using the median-of-ratios method implemented in DESeq2. ",
        "Dispersion estimates were obtained using the default shrinkage estimator. ",
        "A negative binomial generalized linear model was fitted for each gene, ",
        "and the Wald test was used to test for differential expression. ",
        "P-values were adjusted for multiple testing using the Benjamini-Hochberg procedure. ",
        "Genes with an adjusted p-value < 0.05 and an absolute log2 fold change > 1 were considered significantly differentially expressed.\n", sep = "")
  } else if (params$de_algorithm == "edger") {
    cat("Differential expression analysis was performed using edgeR v3.36.0 [@edger_cite]. ",
        "Raw read counts were normalized using the trimmed mean of M-values (TMM) method. ",
        "Dispersion was estimated using the Cox-Reid profile-adjusted likelihood method. ",
        "A negative binomial generalized log-linear model was fitted using the quasi-likelihood framework, ",
        "and the quasi-likelihood F-test was used to identify differentially expressed genes. ",
        "P-values were adjusted for multiple testing using the Benjamini-Hochberg procedure. ",
        "Genes with an adjusted p-value < 0.05 and an absolute log2 fold change > 1 were considered significantly differentially expressed.\n", sep = "")
  } else if (params$de_algorithm == "limma_voom") {
    cat("Differential expression analysis was performed using the limma-voom pipeline [@limma_voom_cite; @voom_cite]. ",
        "Raw counts were first transformed using the voom method, which estimates the mean-variance relationship ",
        "and generates precision weights for each observation. ",
        "TMM normalization was applied prior to voom transformation. ",
        "A linear model was fitted for each gene, and empirical Bayes moderation of standard errors was applied ",
        "using the eBayes function. ",
        "P-values were adjusted for multiple testing using the Benjamini-Hochberg procedure. ",
        "Genes with an adjusted p-value < 0.05 and an absolute log2 fold change > 1 were considered significantly differentially expressed.\n", sep = "")
  }
}
```

`r if(!is.null(params$pathway_analysis) && params$pathway_analysis) "### Pathway Analysis"`

```{r pathway_methods, echo=TRUE, results='asis'}
# Generate pathway analysis methods text
if (!is.null(params$pathway_analysis) && params$pathway_analysis) {
  cat("Pathway analysis was performed using the MITHrIL2 algorithm [@mithril2_cite; @mithril2_method_cite]. ",
      "Pathway topologies were obtained from the Kyoto Encyclopedia of Genes and Genomes (KEGG) database. ",
      "MITHrIL2 extends standard pathway analysis by incorporating post-transcriptional regulatory interactions, ",
      "including microRNA-mediated regulation, to improve the accuracy of pathway impact assessment. ",
      "Pathways with an adjusted p-value < 0.05 were considered significantly impacted.\n", sep = "")
}
```

`r if(!is.null(params$circ_rna_detection) && params$circ_rna_detection) "### Circular RNA Detection"`

```{r circrna_methods, echo=TRUE, results='asis'}
# Generate circular RNA detection methods text
if (!is.null(params$circ_rna_detection) && params$circ_rna_detection) {
  if (!is.null(params$circ_algorithm) && params$circ_algorithm == "ciriquant") {
    cat("Circular RNA detection and quantification were performed using CIRIquant v1.1.2 [@ciriquant_cite], ",
        "which integrates CIRI2 [@ciri2_cite] for back-splice junction identification with pseudo-alignment for accurate circRNA quantification. ",
        "Reads were first aligned to the ", params$genome_name, " reference genome using BWA-MEM [@bwa_cite] ",
        "and HISAT2 [@hisat2_cite] to enable both back-splice junction detection and linear RNA quantification. ",
        "CIRIquant employs a combination of de novo and annotation-assisted approaches to identify and quantify circular RNAs. ",
        "CircRNAs supported by at least 2 back-splice junction reads were retained for downstream analysis.\n", sep = "")
  } else {
    cat("Circular RNA detection was performed using CIRI2 v2.0.6 [@ciri2_cite]. ",
        "Reads were aligned to the ", params$genome_name, " reference genome using BWA-MEM [@bwa_cite] to identify back-splice junctions. ",
        "CIRI2 uses a multiple-seed matching strategy to scan SAM alignment files for paired chiastic clipping (PCC) signals, ",
        "followed by filtering based on GT-AG splicing signals and junction read pairing. ",
        "The ", params$annotation_name, " annotation was used to annotate detected circular RNAs. ",
        "CircRNAs supported by at least 2 back-splice junction reads were retained for downstream analysis.\n", sep = "")
  }
}
```

### Computational Environment

```{r compute_methods, echo=TRUE, results='asis'}
# Document computational resources used
cat("All analyses were performed using ", params$threads, " CPU threads",
    if (!is.null(params$memory_gb)) paste0(" with ", params$memory_gb, " GB of allocated memory") else "",
    ". The pipeline was executed within the RNAdetector containerized environment to ensure reproducibility.\n", sep = "")
```
